---
title: "Assignment 2"
author: "Jocelyn Tsai"
date: "9/13/2020"
output: 
  html_document:
    toc: true
    toc_float: true 
---

I first loaded the tidyverse and ggplot2 libraries:
```{r, message=FALSE}
library(tidyverse)
library(ggplot2)
```

I then loaded my dataset from Assignment 1. Referencing from Catherine's example assignment, I used the "attach" function to connect my dataframe from Assignment 1 to the search path. Now I can refer to my variables without specifying the dataframe that they're in every time.

```{r, message=FALSE}
US_data <- read_csv("wildfire_dataset.csv")
attach(US_data)
```

List of variables from Assignment 1:
(I am not showing "Year" as a variable in this assignment because it does that make sense to graph the distribution of "Year" since there's only 1 value every year from 1929 to 2015)

1. Annual Rainfall in the US (continuous): Total_Rainfall_mm
2. Mean Annual Temperature in the US (continuous): Mean_Temp_Celsius
3. Number of Wildfires in the US each year (continuous): Num_Fires
4. Number of Acres Burned in the US each year (continuous): Acres_Burned
5. Years with Extreme Heat Wave in the US (categorical): Extreme_Heat_Wave
6. Years with High Unemployment in the US (categorical): High_Unemployment

## 1. Statistics in a Table for All Continuous Variables

Statistics include: sample mean, median, standard deviation, inter-quartile range, 95-percent confidence interval for population mean

I used the summary() function to get a quick distribution summary for each of the continuous variables. Since the summary () function does not include standard deviation and it cannot find the 95 percent confidence interval, I also used the sd() and t.test() functions. I then combined the cleaned up statistic values using prettyNum() into a list for each variable.  

### Continuous Variable 1: Annual Rainfall (mm) in the US

```{r}
summary(Total_Rainfall_mm)
rain_summary <- summary(Total_Rainfall_mm)
rain_sd <- sd(Total_Rainfall_mm)
rain_95confidence <- t.test(Total_Rainfall_mm)
```

```{r}
annual_rainfall <- c(prettyNum(rain_summary["Mean"], digits =4),
                     prettyNum(rain_summary["Median"], digits =4),
                     prettyNum(rain_sd, digits = 4),
                     paste(prettyNum(rain_summary["1st Qu."], digits =4), "to",                                             prettyNum(rain_summary["3rd Qu."], digits =4)),
                     paste(prettyNum(rain_95confidence$conf.int[1], digits=4), "to",
                           prettyNum(rain_95confidence$conf.int[2], digits=4)))
annual_rainfall
```


### Continuous Variable 2: Mean Annual Temperature (celsius) in the US

```{r}
summary(Mean_Temp_Celsius)
temp_summary <- summary(Mean_Temp_Celsius)
temp_sd <- sd(Mean_Temp_Celsius)
temp_95confidence <- t.test(Mean_Temp_Celsius)
```

```{r}
mean_annual_temp <- c(prettyNum(temp_summary["Mean"], digits =3),
                      prettyNum(temp_summary["Median"], digits =3),
                      prettyNum(temp_sd, digits = 3),
                      paste(prettyNum(temp_summary["1st Qu."], digits =3), "to",                                             prettyNum(temp_summary["3rd Qu."], digits =3)),
                      paste(prettyNum(temp_95confidence$conf.int[1], digits=3), "to",
                            prettyNum(temp_95confidence$conf.int[2], digits=3)))
mean_annual_temp
```


### Continuous Variable 3: Number of Wildfires in the US each year

```{r}
summary(Num_Fires)
num_fire_summary <- summary(Num_Fires)
num_fire_sd <- sd(Num_Fires)
num_fire_95confidence <- t.test(Num_Fires)
```

```{r}
num_fire <- c(prettyNum(num_fire_summary["Mean"], digits =3),
              prettyNum(num_fire_summary["Median"], digits =3),
              prettyNum(num_fire_sd, digits = 3),
              paste(prettyNum(num_fire_summary["1st Qu."], digits =3), "to",                                         prettyNum(num_fire_summary["3rd Qu."], digits =3)),
              paste(prettyNum(num_fire_95confidence$conf.int[1], digits=3), "to",
                    prettyNum(num_fire_95confidence$conf.int[2], digits=3)))
num_fire
```

### Continuous Variable 4: Number of Acres Burned in the US each year

```{r}
summary(Acres_Burned)
acres_burned_summary <- summary(Acres_Burned)
acres_burned_sd <- sd(Acres_Burned)
acres_burned_95confidence <- t.test(Acres_Burned)
```

```{r}
burned <- c(prettyNum(acres_burned_summary["Mean"], digits =3),
            prettyNum(acres_burned_summary["Median"], digits =3),
            prettyNum(acres_burned_sd, digits = 3),
            paste(prettyNum(acres_burned_summary["1st Qu."], digits =3), "to",                                     prettyNum(acres_burned_summary["3rd Qu."], digits =3)),
            paste(prettyNum(acres_burned_95confidence$conf.int[1], digits=3), "to",
                  prettyNum(acres_burned_95confidence$conf.int[2], digits=3)))
burned
```

### Display all four continuous variables in a formmated table:

```{r}
stats <- c("Sample Mean",
                "Median",
                "Standard Deviation",
                "Interquartile Range",
                "95% Confidence Interval (Population Mean)")
stats
```

```{r}
summary_table <- tibble("Statistics" =stats, 
       "Annual Rainfall (mm)" =annual_rainfall,
       "Mean Annual Temperature (celsius)" =mean_annual_temp,
       "Number of Fires (per year)" =num_fire,
       "Acres Land Burned (per year)" =burned)
summary_table
```


```{r}
knitr::kable(summary_table, caption= "Statistics of Yearly Environmental Data in the US")
```


## 2. Histogram of Distribution for each Continuous Variable

### Continuous Variable 1: Annual Rainfall (mm) in the US

a. Sample Mean: 649.7
b. Median: 651.8
c. Standard Deviation: 41.96
d. Interquartile Range: 627.3 to 676.3	
e. 95% Confidence Interval: 640.8 to 658.7

The distribution of Annual Rainfall is slightly skewed to the left, with a median larger than the sample mean. There is only one outlier (a year) that has extreme rainfall of more than 750 milimeters comparing to other years. 

```{r, message=FALSE}
rain_hist <- ggplot(US_data, aes(x= Total_Rainfall_mm))+
  geom_histogram()

rain_hist+
  theme_light()+
  scale_x_continuous(name= "Annual Rainfall (mm)")+
  scale_y_continuous(name= "Number of Years in Sample")
```


### Continuous Variable 2: Mean Annual Temperature (celsius) in the US

a. Sample Mean: 7.1
b. Median: 7.06
c. Standard Deviation: 0.497
d. Interquartile Range: 6.72 to 7.49
e. 95% Confidence Interval: 7 to 7.21

The distribution of Mean Annual Temperature is slightly skewed to the right, with a mean larger than the median. Values above 7.5 bring the average of the distribution up, while more observations are concentrated on the left side. 

```{r, message=FALSE}
temp_hist <- ggplot(US_data, aes(x= Mean_Temp_Celsius))+
  geom_histogram()

temp_hist+
  theme_light()+
  scale_x_continuous(name= "Mean Annual Temperature (Celsius)")+
  scale_y_continuous(name= "Number of Years in Sample")
```


### Continuous Variable 3: Number of Wildfires in the US each year

a. Sample Mean: 126770
b. Median: 117957
c. Standard Deviation: 55922
d. Interquartile Range: 80075 to 173138
e. 95% Confidence Interval: 114852 to 138689

The distribution of the Number of Wildfires is skewed to the right, with a mean larger than the median. There is a greater number of observations closer to the left side of the histogram. 

```{r, message=FALSE}
fires_hist <- ggplot(US_data, aes(x= Num_Fires))+
  geom_histogram()

fires_hist+
  theme_light()+
  scale_x_continuous(name= "Number of Wildfires")+
  scale_y_continuous(name= "Number of Years in Sample")
```

### Continuous Variable 4: Number of Acres Burned in the US each year

a. Sample Mean: 11477719
b. Median: 5260825
c. Standard Deviation: 13050569
d. Interquartile Range: 3279282 to 14792000
e. 95% Confidence Interval: 8696266 to 14259173

The sample mean is significantly larger than the median, so the histogram is skewed positively to the right. There is a long tail trailing on the right side of the graph, bringing the average of the distribution higher. 

```{r, message=FALSE}
options(scipen = 999)
burned_hist <- ggplot(US_data, aes(x= Acres_Burned))+
  geom_histogram()

burned_hist+
  theme_light()+
  scale_x_continuous(name= "Number of Acres Burned")+
  scale_y_continuous(name= "Number of Years in Sample")
```


## 3. Categorical Variables

### Categorical Variable 1: Years with Extreme Heat Wave

I first used table() to get the number of observations in each category:

```{r}
table(US_data$Extreme_Heat_Wave)
```

I calculated the proportion of the sample (years) that has Extreme Heat Wave:

```{r}
mean(US_data$Extreme_Heat_Wave == "TRUE")
```

I then used the t.test() function to get the 95 percent confidence interval for the proportion that has extreme heat wave.

```{r}
t.test(US_data$Extreme_Heat_Wave == "TRUE")
```

Here is a bar chart showing the number of years that have extreme heat wave versus years without extreme heat wave. 

```{r}
ggplot(US_data, aes(x = Extreme_Heat_Wave, fill= Extreme_Heat_Wave)) +
  geom_bar() +
  scale_y_continuous(name = "Number of Years") +
  scale_x_discrete(name = "Extreme Heat Wave",
                       labels = c("FALSE",
                                  "TRUE")) +
  scale_fill_discrete(name = element_blank(),
                        labels = c("FALSE",
                                   "TRUE")) +
  theme_light() 
```


### Categorical Variable 2: Years with High Unemployment

I first used table() to get the number of observations in each category:

```{r}
table(US_data$High_Unemployment)
```

I calculated the proportion of the sample (years) that has High Unemployment:

```{r}
mean(US_data$High_Unemployment == "TRUE")
```

I then used the t.test() function to get the 95 percent confidence interval for the proportion that has extreme heat wave.

```{r}
t.test(US_data$High_Unemployment == "TRUE")
```

Here is a bar chart showing the number of years that have high unemployment rate versus years that don't have high unemployment rate. 

```{r}
ggplot(US_data, aes(x = High_Unemployment, fill= High_Unemployment)) +
  geom_bar() +
  scale_y_continuous(name = "Number of Years") +
  scale_x_discrete(name = "High Unemployment",
                       labels = c("FALSE",
                                  "TRUE")) +
  scale_fill_discrete(name = element_blank(),
                        labels = c("FALSE",
                                   "TRUE")) +
  theme_light() 
```

